PineAPPL:\\ EW and mixed QCD--EW corrections for PDF fits

### 2020-08-06 paper/sec-results.tex
Clarify one comment

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc MG5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme~\cite{Denner:1999gp,Denner:2005fg} for all massive particles, see
ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that do not involve stable top quarks in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: 
 in particular, PI contributions are consistently included at LO and NLO.\footnote{We employ the $\overline{G}_\mu$ scheme also for the QED coupling entering vertices involving initial-state photons, see section~4.3.3 of ref.\cite{Denner:2019vbn}.}
We use a PDF set that contains a photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast-interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation.

The central values of the renormalisation and factorisation scales, $\mu_\mathrm{R}$ and
$\mu_\mathrm{F}$, are chosen, for each process, as follows. In the case of Drell--Yan
lepton pair production, we use the fixed scale $\mu_\mathrm{R}=\mu_\mathrm{F}=M_\mathrm{Z}$, where $M_\mathrm{Z}$ is
the mass of the Z boson, for the ATLAS measurement, and the scale
$\mu_\mathrm{R}=\mu_\mathrm{F}=M_{\ell\bar\ell}$, where $M_{\ell\bar\ell}$ is the central value of each
invariant mass bin, for the CMS measurement.
In the case of top-quark pair production, we use the dynamic scales
$\mu_\mathrm{R}=\mu_\mathrm{F}=\sqrt{m_\mathrm{t}^2+(p_\mathrm{T}^\mathrm{t})^2}{\Big /}2$ for the distribution differential
in the transverse momentum of the top quark, and $\mu_\mathrm{R}=\mu_\mathrm{F}=H_\mathrm{T}/4$ for the
distribution differential in the invariant mass of the top-quark pair, where
$H_\mathrm{T}=\sqrt{m_\mathrm{t}^2+(p_\mathrm{T}^\mathrm{t})^2}+\sqrt{m_\mathrm{t}^2+(p_\mathrm{T}^{\bar{\mathrm{t}}})}$, with $m_\mathrm{t}$,
$p_\mathrm{T}^\mathrm{t}$ and $p_\mathrm{T}^{\bar{\mathrm{t}}}$ the mass of the top quark and the transverse momenta
of the top and antitop quarks, respectively. These choices were demonstrated
to maximise the convergence of the perturbative expansion~\cite{Czakon:2016dgf}.
In the case of Z-boson production with non-zero transverse momentum, we use
$\mu_\mathrm{R}=\mu_\mathrm{F}=M_\mathrm{Z}$. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted on-the-fly in the Monte Carlo generation upon scale
variations, with the technique presented in Ref.~\cite{Frederix:2011ss}. To this purpose, we use the default \textsc{MG5\_aMC}
implementation, whereby the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and the envelope
from the nine-point scale variations is constructed. However, we note that
\textsc{PineAPPL} allows the user to determine the envelope with any point
prescription, see appendix~\ref{app:pineappl-demo} for an example.

The values of the relevant physical parameters are chosen as
\begin{equation}
\begin{aligned}
M_\mathrm{W} &= \SI{80}{\giga\electronvolt} \text{,} \quad &
M_\mathrm{Z} &= \SI{91.176}{\giga\electronvolt} \text{,} \quad &
m_\mathrm{t} &= \SI{172.5}{\giga\electronvolt} \text{,} \quad \\
\Gamma_\mathrm{W} &= \SI{2.09}{\giga\electronvolt} \text{,} &
\Gamma_\mathrm{Z} &= \SI{2.50}{\giga\electronvolt} \text{,}
\end{aligned}
\label{eq:parameters}
\end{equation}
where $M_\mathrm{W}$, $M_\mathrm{Z}$ and $m_\mathrm{t}$ are the values of the W-boson, Z-boson, and
top quark masses, respectively, and $\Gamma_\mathrm{W}$ and $\Gamma_\mathrm{Z}$ are the width of
the W boson and of the Z boson, respectively. The value of the strong
coupling is chosen consistently with the PDF set, $\alphas(M_\mathrm{Z})=0.118$.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon distance. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematics observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination wouldl have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections,
see refs.~\cite{Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order as the EW corrections, they are not included in our
computations. In fact, while nothing prevents to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of ``standard'' EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases {\bf MZ OTHERS???}).

Finally, we implement the kinematic cuts specified in the corresponding
experimental analyses. In the case of high-mass Drell--Yan lepton pair
production at \SI{7}{\tera\electronvolt} measured by ATLAS, we require $p_\mathrm{T}^\ell>\SI{25}{\giga\electronvolt}$,
$|\eta_\ell|<2.5$ and $\SI{116}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$ for the transverse
momentum and the pseudorapidity of each lepton and for the invariant mass of the
lepton pair, respectively. In the case of double-differential Drell--Yan
lepton-pair production
at \SI{7}{\tera\electronvolt} measured by CMS, we require $p_\mathrm{T}^{\ell_1}>\SI{14}{\giga\electronvolt}$, $p_\mathrm{T}^{\ell_2}>\SI{9}{\giga\electronvolt}$,
$|\eta_\ell|<2.4$, $|\eta_{\ell\bar\ell}|<2.4$ and $\SI{20}{\giga\electronvolt}<M_{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$
for the transverse momentum and the pseudorapidity of each lepton, and for the
pseudorapidity and the invariant mass of the lepton pair. In the case of Z-boson
production with non-zero transverse momentum at \SI{13}{\tera\electronvolt} measured by CMS,
require $p_\mathrm{T}^\ell>\SI{25}{\giga\electronvolt}$, $|\eta_\ell|<2.4$,
$M_\mathrm{Z}-\SI{15}{\giga\electronvolt}<M_{\ell\bar\ell}<M_\mathrm{Z}+\SI{20}{\giga\electronvolt}$,
$|\eta_{\ell\bar\ell}|<2.4$ and $\SI{20}{\giga\electronvolt}<p_\mathrm{T}^{\ell\bar\ell}<\SI{1500}{\giga\electronvolt}$ for the
transverse momentum and pseudorapidity of each lepton, and for the invariant mass,
pseudorapidity and transverse momentum of the lepton pair. Imposing a minimum cut for the transverse momentum of the lepton pair 
is necessary as we stick to a fixed-order computation, without including resummation effects.

\subsection{Numerical results}
\label{subsec:numerical_results}

For each of the measurements discussed in
section~\ref{subsec:processes_and_measurements}, we compute the expectation
value of the corresponding observable for each kinematic bin in two different
ways: directly, by means of \textsc{MG5\_aMC}, and a posteriori, by convolving
the fast-interpolation grid produced by \textsc{PineAPPL} with the PDF set
specified in section~\ref{subsec:computational_settings}. In the following, we
will refer to the first result as the MC result, and to the second as
the \textsc{PineAPPL} result. We repeat the computation for theories accurate
to NLO QCD and to NLO QCD+EW, respectively. The corresponding orders of the
strong and electroweak couplings that we consider have been specified in
section~\ref{subsec:processes_and_measurements}. In each case, we determine the
PDF uncertainty (coming from the PDF ensemble), the scale uncertainty (coming
from variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of a fraction of per mille on the total cross section. While
this choice does not affect the validation of the \textsc{PineAPPL} result
against the MC result, it ensures that the statistical uncertainty of
the computation remains negligible in comparison to the PDF and scale
uncertainties, as we will explicitly demonstrate. This is a desirable feature
to correctly interpret the size of the EW corrections.

@@

The process receives its leading contribution from the $\mathrm{gg}$ channel,
which varies between 81\% and 61\% (76\% and 83\%) of the $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$) differential cross section as the value of the
transverse momentum of the top quark (the invariant mass of the top-quark pair)
increases in the bin range; $\mathrm{\gamma}\mathrm{g}$ scattering
correspondingly accounts for about 0.5\% -- 1\% (0.5\% -- 0.7\%) of the cross
section, and is almost entirely {\bf MZ CHECK!!!} due to the LO contribution at $\mathcal{O}(\alphas\alpha)$; the contribution from other PI parton luminosities is comparatively
negligible. Overall, the EW corrections suppress the $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$) distribution by about 0.2\% -- 3.5\%
(0.5\% -- 0.2\%) for increasing values of $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$), except in the first bin of the
$p_\mathrm{T}^\mathrm{t}$ distribution, where they enhance the cross section by
about 1\%. The size of these shifts, however, remains always significantly
smaller than the data uncertainty.\footnote{In
  figure~\ref{fig:atlastop} the data uncertainty corresponds to the ATLAS
  measurement~\cite{Aad:2015auj}. Similar considerations apply also for the CMS
  measurement~\cite{Khachatryan:2015oaa}.}
As a consequence, we anticipate that the more accurate NLO QCD+EW theory is
likely to be easily accommodated by the large data uncertainty, should the data
be fitted with the inclusion of EW corrections; PDFs will possibly not become
more precise.

>>>>

We generate each process by means of the Universal FeynRules Output
(UFO)~\cite{Degrande:2011ua} model {\tt loop\_qcd\_qed\_sm\_Gmu},
included as standard in {\sc MG5\_aMC}. It contains the UV and $R_2$
counterterms relevant to NLO QCD and EW corrections, the latter in the
$\overline{G}_\mu$ scheme. The model features five massless quark flavours,
sets the CKM matrix equal to the identity, and is compatible with the usage of
the complex mass (CM) scheme~\cite{Denner:1999gp,Denner:2005fg} for all massive particles, see
ref.~\cite{Frederix:2018nkq} for details. We use this scheme
for all processes that involve only massless particles in the final state.
The photon is always considered as part of the proton in the initial state and
of any hadronic jet produced in the final state: 
 in particular, PI contributions are consistently included at LO and NLO.\footnote{We employ the $\overline{G}_\mu$ scheme also for the QED coupling entering vertices involving initial-state photons, see section~4.3.3 of ref.\cite{Denner:2019vbn}.}
We use a PDF set that contains a photon PDF, namely
{\tt NNPDF31\_nlo\_as\_0118\_luxqed}~\cite{Bertone:2017bme}. We evaluate the PDF
uncertainty associated to the theoretical predictions a posteriori, that is,
we convolve the fast-interpolation grid generated with {\sc PineAPPL} with
each member in the PDF set, and we compute the associated standard deviation.

The central values of the renormalisation and factorisation scales, $\mu_\mathrm{R}$ and
$\mu_\mathrm{F}$, are chosen in a process specific way, and are discussed in sec.~\ref{subsec:processes_and_measurements}. In order to estimate the missing higher-order uncertainty,
we allow the events to be reweighted on-the-fly in the Monte Carlo generation upon scale
variations, with the technique presented in Ref.~\cite{Frederix:2011ss}. To this purpose, we use the default \textsc{MG5\_aMC}
implementation, whereby the factorisation and renormalisation scales
are varied down to a factor $1/2$ and up to a factor $2$, and the envelope
from the nine-point scale variations is constructed. However, we note that
\textsc{PineAPPL} allows the user to determine the envelope with any point
prescription, see appendix~\ref{app:pineappl-demo} for an example.

The values of the relevant physical parameters are chosen as
\begin{equation}
\begin{aligned}
M_\mathrm{W} &= \SI{80}{\giga\electronvolt} \text{,} \quad &
M_\mathrm{Z} &= \SI{91.176}{\giga\electronvolt} \text{,} \quad &
m_\mathrm{t} &= \SI{172.5}{\giga\electronvolt} \text{,} \quad \\
\Gamma_\mathrm{W} &= \SI{2.09}{\giga\electronvolt} \text{,} &
\Gamma_\mathrm{Z} &= \SI{2.50}{\giga\electronvolt} \text{,}
\end{aligned}
\label{eq:parameters}
\end{equation}
where $M_\mathrm{W}$, $M_\mathrm{Z}$ and $m_\mathrm{t}$ are the values of the W-boson, Z-boson, and
top quark masses, respectively, and $\Gamma_\mathrm{W}$ and $\Gamma_\mathrm{Z}$ are the width of
the W boson and of the Z boson, respectively. The value of the strong
coupling is chosen consistently with the PDF set, $\alphas(M_\mathrm{Z})=0.118$.
The definition of observables and cuts is process-specific, and it follows the corresponding experimental measurements.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon distance. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematics observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination would have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections,
see refs.~\cite{Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order as the EW corrections, they are not included in our
computations. In fact, while nothing prevents to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of ``standard'' EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases {\bf MZ OTHERS???}).

\subsection{Results for specific processes and measurements}
\label{subsec:processes_and_measurements}

We focus on the following three processes: Drell--Yan lepton-pair production,
top-quark pair production, and Z-boson (lepton-pair) production with non-zero
transverse momentum in proton-proton collisions. For each of these processes,
we present the experimental measurements, process-specific settings, and phenomenological results.

When presenting the results, for each of the processes and measurements considered, we compute the expectation
value of the corresponding observable for each kinematic bin in two different
ways: directly, by means of \textsc{MG5\_aMC}, and a posteriori, by convolving
the fast-interpolation grid produced by \textsc{PineAPPL} with the PDF set
specified in section~\ref{subsec:computational_settings}. We
will refer to the first result as the MC result, and to the second as
the \textsc{PineAPPL} result. We repeat the computation for theories accurate
to NLO QCD and to NLO QCD+EW, respectively. The corresponding orders of the
strong and electroweak couplings that we consider are specified for each process. In each case, we determine the
PDF uncertainty (coming from the PDF ensemble), the scale uncertainty (coming
from variations of the factorisation and renormalisation scales), and the Monte
Carlo uncertainty (coming from the finite number of events generated). In this
last respect, we consider by default high-statistics computations, whereby we
require a relative Monte Carlo precision of a fraction of per mille on the total cross section. While
this choice does not affect the validation of the \textsc{PineAPPL} result
against the MC result, it ensures that the statistical uncertainty of
the computation remains negligible in comparison to the PDF and scale
uncertainties, as we will explicitly demonstrate. This is a desirable feature
to correctly interpret the size of the EW corrections.

@@

The process receives its leading contribution from the $\mathrm{gg}$ channel,
which varies between 81\% and 61\% (76\% and 83\%) of the $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$) differential cross section as the value of the
transverse momentum of the top quark (the invariant mass of the top-quark pair)
increases; the largest PI contribution for this process is $\gamma\mathrm{g}$ scattering,
which accounts for about 0.5\% -- 1\% (0.5\% -- 0.7\%) of the cross
section, and is almost entirely (90\%) due to the LO contribution at $\mathcal{O}(\alphas\alpha)$; the contribution from other PI parton luminosities is comparatively
negligible. Overall, the EW corrections suppress the $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$) distribution by about 0.2\% -- 3.5\%
(0.5\% -- 0.2\%) for increasing values of $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$), except in the first bin of the
$p_\mathrm{T}^\mathrm{t}$ distribution, where they enhance the cross section by
about 1\%. The size of these shifts, however, remains always significantly
smaller than the data uncertainty.\footnote{In
  figure~\ref{fig:atlastop} the data uncertainty corresponds to the ATLAS
  measurement~\cite{Aad:2015auj}. Similar considerations apply also for the CMS
  measurement~\cite{Khachatryan:2015oaa}.}
As a consequence, we anticipate that the more accurate NLO QCD+EW theory is
likely to be easily accommodated by the large data uncertainty, should the data
be fitted with the inclusion of EW corrections; PDFs will possibly not become
more precise.
